\documentclass[10pt,a4paper,oneside]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{enumerate}
\date{June 24, 2019}
\author{Baboo J. Cui, Yangang Cao}
\title{Properties Q\&A}
\begin{document}
\maketitle
\tableofcontents

\newpage
\section{Basic}
\begin{enumerate}[1.]
\item    How to see all branches in git?   \\
Use command \$ git branch
\item    How to set global user info?   \\
Use command \$ git config --global - -edit
\item    How to set user name?    \\
Use command \$ git config - -global user.name “NAME”
\item   How to set user e-mail?    \\
Use command \$ git config - -global user.email “EMAIL\_ADD”
\item    How to install brew?   \\
Use command \$ /usr/bin/ruby -e ``\$(curl -fsSL https://raw.githubusercontent.com\\/Homebrew/install/master/install)''
\item    How to install bash-completion?   \\
1.Use command \$ brew install bash-completion\\ 2.Add [[ -r ``/usr/local/etc/profile.d/bash\_completion.sh'' ]] \&\& . ``/usr/local\\/etc/profile.d/bash\_completion.sh''  to ./bash\_profile \\3.Then activate it by ~/.bash\_profile
\item  How to create a new branch?     \\
Use command \$ git checkout -b “NEW\_BRANCH\_NAME”
\item   How to move to a branch?    \\
Use command \$ git checkout “EXIST\_BRANCH\_NAME”
\item   How to track an origin branch?    \\
Use command \$ git branch -t LOCAL\_TRACKING\_NAME origin/REMOTE\_NAME (Don't create local that already exits)
\item    How to push for the first time?   \\
Use command \$ git push ORIGIN BC\_DEV, must be very specific
\item    How to pull for the first time?   \\
Use command \$ git pull ORIGIN MASTER, must be very specific
\end{enumerate}
\end{document}