\documentclass[10pt,a4paper,oneside]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{enumerate}
\date{June 21, 2019}
\author{Baboo J. Cui, Yangang Cao}
\title{HTML Q\&A}
\begin{document}
\maketitle
\tableofcontents

\newpage
\section{Basic}
\begin{enumerate}[1.]
\item  What is HTML?   \\
Hypertext markup language, a text file corresponds to a web page
\item   The two main parts of HTML?  \\
<head> and <body>, meta info should be in <head> and content info should be in <body>, both should be within <html>
\item    How HTML is declared? \\
<html> \dots </html>, all info should be inside of it
\item  What a web page essentially is?   \\
A text file
\item  Static page vs dynamic page?   \\
Static one can be directly viewed in browser and dynamic one needs to be parsed
\item  What is tag?   \\
Contents in <>, like <head>, <style>
\item  What is the range of a tag?   \\
Between the dual tag, like <body>…</body>, a start one and an end one
\item  What does start tag includes?   \\
The element name and its optional properties, both must be in start tag
\item  What does element properties includes?   \\
Property name and value(optional, should be in ``\dots''), if more than one, separated by white space <tag a=``v1'' b=``v2''>\dots</tag>
\item   What is empty tag?  \\
Not in dual ones, only start tag, no end one. It can contains properties
\item    What is tag nesting? \\
A tag content can be included in the other
\item   HTML structural integrity?  \\
4 parts: Doc type<!DOCTYPE\dots>, namespace<html xmlns=``\dots''>, coding<meta charset=``\dots''> and title<title>\dots</title>
\item  Comments for code?   \\
Put comments in <!--XXX-->
\item  Extension name of HTML?   \\
.htm or .html
\item  How title is declared?   \\
\end{enumerate}
\end{document}