\documentclass[10pt,a4paper,oneside]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{enumerate}
\date{June 21, 2019}
\author{Baboo J. Cui, Yangang Cao}
\title{Vue Q\&A}
\begin{document}
\maketitle
\tableofcontents

\newpage
\section{Preparation}
\begin{enumerate}[1.]
\item  What is Vue?  \\
A framework for frontend development
\item What is node?   \\
A javascript environment
\item   Env requirement? \\
List: brew, node, yarn, npm, node is language that vue need, the rest for package management
\item  How to install npm in centOS?  \\
Use command \$ yum install npm, this will also install the dependency nodejs
\item How to install yarn in centOS?   \\
Check the net for repository
\item   Where does npm or yarn installation goes? \\
To path: /usr/local/lib/node\_modules/MODULE\_NAME
\item   Advantages of yarn? \\
It’s fast, yarn is preferred than npm
\item  Name of vue client?  \\
The name is @vue/cli
\item  What is @vue/cli?  \\
A system for development based on Vue.js
\item  How to install @vue/cli?  \\
Use command \$ npm install -g @vue/cli or \$ yarn global add @vue/cli, and vue command now is available
\item  How to check if vue is installed correctly?  \\
Use command \$ vue - -version
\item   Npm source optimization? \\
Use command \$ npm config set registry https://registry.npm.taobao.org and use \$ npm config get registry to check
\item   Yarn source optimization?  \\
Use command \$ yarn config set registry https://registry.npm.taobao.org and use \$ yarn config get registry to check
\item  How to install dependency for a project?  \\
Go to the project folder and use command \$ yarn install, this will automatically install all necessary packages in packages.json
\item   Common framework for frontend? \\
Vue, React, Angular…
\end{enumerate}

\section{Fast Prototype (Just know this!)}
\begin{enumerate}[1.]
\item  How to achieve fast prototype dev?  \\
Use \$ vue serve and \$ vue build, this requires @vue/cli-service-global
\item  Disadvantages of vue serve?  \\
Must be global, which may cause compatible problem
\item  How to install @vue/cli-service-global?  \\
Use command \$ npm install -g @vue/cli-service-global or \$ yarn global add @vue/cli-service-global
\end{enumerate}

\section{Get Started}
\begin{enumerate}[1.]
\item  How to create a project?  \\
Go to target folder and use command \$ vue create PROJECT\_NAME, and config it by indication
\item  How to create a project with GUI?  \\
Use command \$ vue ui, this will start a webpage for creating a project
\item  Any common options for vue create?  \\
-n for no git, -f for overwrite config file, -x use proxy, etc. Just remember these 3
\item  How to start a project server?  \\
Go to folder and use command \$ yarn serve, and visit http://localhost:8080/, you can see the page
\item  What frame does @vue/cli use?  \\
Based on plugins, dependencies are start with @vue/cli-plugin-, most of characters are achieved by plugins
\item   How to manage plugins? \\
Use vue GUI or use command \$ vue add PLUGIN\_NAME
\item What is the purpose of \$ vue add?   \\
Install and use vue client plugins, NOT for replacing npm packages, essentially it call npm to to install
\item  What is the purpose o adding @vue prefix?  \\
For control the scope, If without @vue, the command will parse it as unscoped package
\item  How to add config outside of the project?  \\
In package.json, add FOLDER\_PATH to vuePlugins.resolveFrom
\item  How to visit API of plugins instead of creating one?  \\
In package.json, add vuePlugins.service entry, note all plugins have API
\item  What is vue cli preset?   \\
A json file to control predefined preferences option and plugins
\item  Where is vue cli preset?  \\
At path: ~/.vuerc, this is configurable
\item  How to run command?  \\
Use command \$ npm run COMMAND or \$ yarn COMMAND, all commands are in scripts in package.json
\item   How to use serve command? \\
Vue-cli-service serve, to start service for development with hot-module-replacement. Directly use command \$ yarn serve
\item  How to use build command?  \\
Vue-cli-service build, create file at dist/ for production. Directly use \$ yarn build
\item  How to use inspect command?  \\
To inspect internal webpack config
\item   How to see all possible command? \\
Use command \$ npx vue-cli-service help
\item  What is npx?  \\

\item How to clear catch to avoid possible bugs?   \\
Delete everything in node\_modules/.cache
\item   How to run /dist file directly? \\
After install serve command, use command \$ serve -s dist, where [-s] means single-page application	
\end{enumerate}
\end{document}